Gusto: a toolkit for compatible finite element dynamical cores

### 2025-08-05 source/paper.tex
a bit more typing about transport

\subsubsection{Transport of $\mathbb{V}^{HDiv}$ fields}

These schemes do not preserve the monotonicity of the continuous
equation, generating spurious over and undershoots. When monotonicity
is important, for example, to avoid generating spurious and unphysical
negative values of moisture variables, we can apply a limiter as part
of the transport scheme.

\JScomment{I think that we should at least mention that we can do horizonal and vertical splitting for stability - maybe don't need to write out all of the forms though!}

>>>>

\subsubsection{Transport of $\mathbb{V}^{HDiv}$ fields}
All three of the equation sets described in section \ref{} include a
prognostic equation for the velocity field
$\MM{u}\in\mathbb{V}^{HDiv}$ that contains a nonlinear transport term
$\MM{u}\cdot\nabla\MM{u}$. This term can be discretised directly using
a vector version of the scheme in section ** or, as in \citet{} it can
be replaced by the vector invariant form and then discretised. In
Gusto we offer both options and we present the vector invariant scheme
below.

The schemes described above do not preserve monotonicity unlike the
continuous equation, generating spurious over and undershoots. When
monotonicity is important, for example, to avoid generating spurious
and unphysical negative values of moisture variables, we can apply a
limiter as part of the transport scheme. These limiters are described
in \ref{}.

It is common in weather and climate models for the horizontal and
vertical directions to be discretised differently due to the different
scales of motion in each direction. Typically the vertical part of the
transport term is treated implicitly so that the timestep is not
restricted by the vertical Courant number. We provide the option to
split the transport terms into horizontal and vertical components, as
described in \ref{}.
